---
title: "Maps Visuals"
author: "Tonatiuh De Leon"
output: 
  html_document:
    theme: flatly
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

This section of my portfolio contains many **Maps Visualizations** I've graphed throughout the years. Take a look at them! I created each graph from scratch in R, check out the code with the `show` buttons to the top right of each graph.

# {.tabset .tabset-fade .tabset-pills}

## Kwh/Capita Static

```{r, fig.width=10, fig.width=10, message=FALSE, warning=FALSE}

library(tidyverse)
library(sf)
library(viridis)
library(RColorBrewer)
library(cowplot)

enuse <- read_csv("enuse.csv") %>%
  rename(enused = `Primary energy consumption per capita (kWh/person)`, region = Entity, year = Year, code = Code) %>% 
  filter(year == 2021) %>% 
  mutate(region = case_when(
  region == "United States" ~ "USA",
  region == "Democratic Republic of Congo" ~ "Democratic Republic of the Congo" ,
  region == "Congo" ~ "Republic of Congo",
  region == "Cote d'Ivoire" ~ "Ivory Coast",
  region == "United Kingdom" ~ "UK",
  TRUE ~ region))


enuse$enused <- log(enuse$enused)

breaks <- c(0, 1, 2, 3, 4, 5)
labels <- as.character(breaks)

world_map <- map_data("world") %>% 
  filter(!(region %in% c("French Southern and Antarctic Lands", "Antarctica")))

world_map <- left_join(world_map, enuse, by = "region")

ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = enused), color = "grey25")+
  
  scale_fill_gradientn(aesthetics = "fill",
                       colors = c(
                         "#e0f3dd", 
                         "#f0f0d9", 
                         "#fee090", 
                         "#fdae61", 
                         "#f46d43", 
                         "#a50026"),
                       
                       limits=c(5,12.3),
                       labels = c("0 kWh", "1,000 kWh", "3,000 kWh", "10,000 kWh", "30,000 kWh", "100,000 kWh"),
                       breaks = c(5,6,8,9.5,11,12.3),
                       guide = guide_colorbar(
                         barwidth = 20, 
                         barheight = 0.5)) +
  labs(
    fill = "kWh Per Capita"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size=12, color = "grey5"),
    legend.title.position = "top",
    legend.text = element_text(size = 9, color = "grey25", vjust = 3)
  ) 

```

This map portrays the energy used in Kw/h per Capita Worldwide. Darker colored countries are the ones with the highest rating.

<br>

## Kwh/Capita Interactive

```{r message=FALSE, warning=FALSE, fig.align='center'}
library(tidyverse)
library(leaflet)

enuse <- read_csv("enuse.csv") |> 
  rename(enused = `Primary energy consumption per capita (kWh/person)`, 
         region = Entity, 
         year = Year, 
         code = Code ) |> 
  filter(year == 2021 ) |> 
  mutate(region = case_when(
  #region == "United States" ~ "USA",
  region == "Democratic Republic of Congo" ~ "Democratic Republic of the Congo" ,
  region == "Congo" ~ "Republic of Congo",
  region == "Cote d'Ivoire" ~ "Ivory Coast",
  TRUE ~ region )) |> 
  mutate(enused = enused / 1000 ) |> 
  drop_na()

coord <- read_csv("coord.csv")|> 
  inner_join(enuse, by="region")


coord$enused <- round(coord$enused, 2)

write_csv(coord, "coords.csv")


leaflet(coord) |> 
  addTiles() |> 
  addMarkers(lat=~lat, lng= ~long, 
             popup = ~paste("Region:", 
                            region, "<br>", 
                            "Energy consumption:", 
                            enused, "thousands of Kw/h per capita"))

```

This is the same map as the first one, difference is the style and formatting. I prefer the first one, but both have their pros.

<br>

## U.S. Most Dense Cities

```{r fig.height=10, fig.width=12, fig.width=10, message=FALSE, warning=FALSE}
library(tidyverse)
library(USAboundaries)
library(dplyr)
library(sf)
library(USAboundariesData)
library(ggrepel)

idahoc <- us_counties(states = "Idaho")


cities <- us_cities() %>% 
  
  filter(
    
    state_name != "Alaska", 
    
    state_name != "Hawaii", 
    
    state_name != "District of Columbia", 
    
    state_name != "Puerto Rico")


contstate <- us_states() %>% 
  
  filter(name != "Alaska", 
         
         name != "Hawaii",
         
         name != "District of Columbia", 
         
         name != "Puerto Rico")


tops <- cities %>% 
  
  group_by(state_name) %>% 
  
  mutate(rank = round(dense_rank(desc(population)))) %>% 
  
  ungroup() %>% 
  
  filter(rank <= 3)


top1 <- tops %>% 
  
  filter(rank ==1) 

ggplot() +

  geom_sf(data = contstate, fill = NA, color = "grey50") +  
  geom_sf(data = tops, mapping = aes(size = population, color = rank), alpha = 0.8) +  
  

  geom_label_repel(
    data = subset(tops, rank == 1),
    mapping = aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    color = "darkblue",
    size = 5,
    max.overlaps = 50,
    fontface = "bold"  
  ) +

  guides(
    size = guide_legend(title = "Population (Thousands)", 
                        override.aes = list(color = "grey50")), 
    color = guide_legend(title = "City Rank", 
                         override.aes = list(size = 5))  
  ) +


  scale_color_continuous(
    breaks = c(1, 2, 3),  
    limits = c(1, 3)      
  ) +
  
  scale_size_continuous(
    breaks = seq(20000, 800000, by=200000), 
    labels = c("200k","400k","600k", "800k" ), 
    limits = c(200000, 800000)      
  ) +


  labs(
    title = "Most Densely Populated Cities in the U.S.",
    subtitle = "Population sizes represented by point size | Population density rank by color (1 most dense)",
    x = NULL,  
    y = NULL   
  ) +
  

  theme_minimal() +
  theme(
    panel.grid = element_blank(),   
    axis.text = element_blank(),    
    axis.ticks = element_blank(),   
    legend.position = "bottom",     
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold", color = "grey5"),
    plot.subtitle = element_text(hjust = 0.5, size = 16, color = "grey30"),
    legend.title = element_text(size = 15, color = "grey15"),
    legend.text = element_text(size = 13, color = "grey25"),
  )



```

What cities in the U.S. have the highest population density overall?

<br>

## Idaho's Water System

```{r, fig.width=10, message=FALSE, warning=FALSE}

library(tidyverse)

library(sf)

library(ggrepel)

library(USAboundaries)

# Read in all four of the necessary SHP file datasets.

pacman::p_load(downloader, sf, fs, tidyverse)

wellspath <- "https://byuistats.github.io/M335/data/Wells.zip"

df <- tempfile(); uf <- tempfile()

download(wellspath, df, mode = "wb")

unzip(df, exdir = uf)

wells <- read_sf(uf)

file_delete(df); dir_delete(uf)

damspath <- "https://byuistats.github.io/M335/data/Idaho_Dams.zip"

df2 <- tempfile(); uf2 <- tempfile()

download(damspath, df2, mode = "wb")

unzip(df2, exdir = uf2)

dams <- read_sf(uf2)

file_delete(df2); dir_delete(uf2)

waterpath <- "https://byuistats.github.io/M335/data/water.zip"

df3 <- tempfile(); uf3 <- tempfile()

download(waterpath, df3, mode = "wb")

unzip(df3, exdir = uf3)

water <- read_sf(uf3)

file_delete(df3); dir_delete(uf3)

stateshapepath <- "https://byuistats.github.io/M335/data/shp.zip"

df4 <- tempfile(); uf4 <- tempfile()

download(stateshapepath, df4, mode = "wb")

unzip(df4, exdir = uf4)

stateshape <- read_sf(uf4)

file_delete(df4); dir_delete(uf4)


ID_counties <- USAboundaries::us_counties(states = "ID")

wellsfiltered <- wells |> 

group_by(WellID) |> 

filter(Production >= 5000)

damsfiltered <- dams |>

filter(SurfaceAre >= 50)

waterfiltered <- water |>

filter(FEAT_NAME %in% c('Snake River', 'Henrys Fork'))


ggplot() +

  geom_sf(data = ID_counties, 
          fill = "grey99") +
  
  geom_sf(data = damsfiltered, 
          aes(color = "Dams"), 
          show.legend = TRUE) +
  
  geom_sf(data = waterfiltered, 
          aes(color = "Rivers"), 
          linewidth = 1, 
          show.legend = TRUE) +
  
  geom_sf(data = wellsfiltered, 
          aes(color = "Wells"), 
          show.legend = TRUE) +
  
  theme_bw() +
  
  coord_sf(crs = st_crs(4267)) +
  
  scale_color_manual(
    name = "Features",
    values = c(
      "Rivers" = "steelblue1", 
      "Dams" = "chocolate", 
      "Wells" = "darkseagreen4"),
    labels = c("Rivers", 
               "Dams", 
               "Wells")) +
  
  theme_minimal() +
  
  labs(
    title = "Idaho's Dams, Wells, and Rivers"
  ) +
  
  theme(
    
    panel.grid = element_blank(), 
    
    axis.text = element_blank(),  
    
    axis.ticks = element_blank(),  
    
    legend.position = "right",  
    
    plot.title = element_text(hjust = 0.5, 
                              size = 18, 
                              face = "bold", 
                              color = "grey5"),
    
    plot.subtitle = element_text(hjust = 0.5, 
                                 size = 16, 
                                 color = "grey30"),
    
    legend.title = element_text(size = 12, 
                                color = "grey15"),
    
    legend.text = element_text(size = 11, 
                               color = "grey25")) +
  
  guides(
    color = guide_legend(override.aes = list(size = 4, 
                                             shape = 16, 
                                             linetype = NA)))

  
```

A summary of Idaho's key rivers, dams, and wells. Overall Idaho's water system portrayed in a map visualization.

<br>

