---
title: "Finding an Apartment in Rexburg"
author: "Tonatiuh De Leon"
output: 
  html_document:
    theme: flatly
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r, message=FALSE, warning=FALSE}
library(mosaic)
library(tidyverse)
library(plotly)
library(pander)
library(DT) 
library(scales)
library(formattable)
            # If you get an error stating: 
            # Error in library(DT): there is no package called 'DT'
            # You will need to run: install.packages("DT") 
            # in your Console, then try "Knit HTML" again.
Rent <- read_csv("Rent.csv")

Steph_apt <- Rent %>% filter(AvgFloorPlanCost/4 < 300 & Residents > 75 & Gender == "F") %>%
  mutate(Price = AvgFloorPlanCost/4)

StephanieRent <- Rent |> 
  filter(
    Gender=="F",
    AvgFloorPlanCost < 1050,
    Residents > 140,
    CrowFlyMetersToMC < 630)

```

<br>

# {.tabset .tabset-fade .tabset-pills}

## Background

---

A friend wanted to learn about what housing options she had for living at BYU-Idaho. She was trying to find something that is close to campus and around $300 a month in rent. I'm good making friends with roommates, but I would like somewhere that has a lot of people around so I can get to know as many people as possible.

I'll use a dataset provided at my school with housing data on Rexburg to find out my best options.

---

**Toggle between tabs to see the analysis**


## Introduction

<br>

Friend, in order to analyze all the women apartments in Rexburg. We will consider how they are distributed by price, number of students, and closeness to Campus. Play Around with the graph below. It has all Women apartments in Rexburg. Then continue reading to find the best option for you

<br>

```{r, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
Housing <- Rent %>% 
  filter(Gender == 'F') %>% 
  mutate(CrowFlyMetersToCenter = round((CrowFlyMetersToMC + CrowFlyMetersToRicks)/2)) %>% 
  filter(CrowFlyMetersToCenter <= 805) %>% 
  mutate(AvgCostByMonth = round(AvgFloorPlanCost/3.22)) %>% 
  filter(AvgCostByMonth <= 350) %>% 
  rename(Apartment = Name) %>%
  select(Apartment, Residents, ParkingSpaces, Deposit, CrowFlyMetersToCenter, AvgCostByMonth)  

#favstats(Housing$CrowFlyMetersToCenter)

HousingTable <- Housing %>% 
  mutate(distance = paste('Distance:', CrowFlyMetersToCenter, '\n')) %>% 
  mutate(MRent = paste('MonthlyRent:', AvgCostByMonth, '\n')) %>%
  mutate(RCount = paste('ResidentCount:', Residents)) %>%
  mutate(apt = paste(Apartment, '\n')) %>% 
  mutate(HData = paste(apt, distance, MRent, RCount))
# 
# plot_ly(
#   HousingTable,
#   y= ~AvgCostByMonth, 
#   x= ~CrowFlyMetersToCenter, 
#   color= ~-Residents, 
#   colorscale='viridis',
#   hovertemplate = paste(
#     "<b>Housing Information</b><br>",
#     "Apartment: ", HousingTable$Apartment, "<br>",
#     "Distance To Campus: ", HousingTable$CrowFlyMetersToCenter, " meters<br>",
#     "Monthly Rent: ", HousingTable$AvgCostByMonth, "<br>",
#     "Resident Count: ", HousingTable$Residents
#   ),
#   hoverinfo = "text"
# ) %>%
# layout(
#   title="Housing Options within Stephanie's Criteria",
#   xaxis=list(title="Distance from the center of Campus (meters)"),
#   yaxis=list(title="Average Cost per Month")
# )


Rent <- Rent |> 
  filter(Gender != "M")

plot_ly(
  data = Rent, 
  x = ~AvgFloorPlanCost, 
  y = ~Residents, 
  z = ~CrowFlyMetersToMC, 
  type = "scatter3d", 
  mode = "markers", 
  marker = list(color = "steelblue"), 
  hovertemplate = paste(
    'Apartment complex: ', Rent$Name, '<br>',
    'Price: ', dollar(Rent$AvgFloorPlanCost), '<br>',
    'Students: ', Rent$Residents, '<br>',
    'Distance: ', round(Rent$CrowFlyMetersToMC)),
  color = I("steelblue")) %>%
  layout(
    title = "Rexburg Women's Apartments Comparison",
    scene = list(
      xaxis = list(title = "Price"),
      yaxis = list(title = "Capacity"),
      zaxis = list(title = "Walking Minutes"))
  )


```

<br><br>

## Statistical Description

<br>

To continue finding the best option for you, here is table with the statistics of all Rexburg apartments so that you can see what the mean is and see a more realistic approach to your needs.

```{r,echo=F, warning=F, message = F}

# plot_ly(
#   data = StephanieRent, 
#   x=~AvgFloorPlanCost, 
#   y=~Residents, 
#   marker=list(color="steelblue"), 
#   text=paste0(
#     'Apartment complex: ', StephanieRent$Name, 
#     '# Price: ', StephanieRent$AvgFloorPlanCost,
#     ' Students: ', StephanieRent$Residents)) %>% 
#   layout(
#     title="Women's Apartments Comparisson", 
#     xaxis=list(title="Apartment price per semester"), 
#     yaxis=list(title="Number Of Students In Apartment Complex"))

dt <- Rent |> 
  filter(Gender == "F")

rbind(
  Price = favstats(dt$AvgFloorPlanCost)[c("min", "mean", "max")],
  Capacity = favstats(dt$Residents)[c("min", "mean", "max")],
  'Walking Minutes' = favstats(dt$CrowFlyMetersToMC)[c("min", "mean", "max")]) |> pander()

```

<br><br>

## Analysis on Price vs. Residents

<br>

Now, let us only consider those apartments with a price between 800 and 1050, also all those who are below the mean in walking time to campus, and finally those who are above the mean in resident capacity.

<br>

```{r, fig.align='center', warning=F, message=F}
ggplot(Rent, aes(AvgFloorPlanCost , Residents)) +
  geom_point(col="steelblue", alpha=.5)+
  ggrepel::geom_label_repel(data=StephanieRent, aes(label = Name), size = 3, nudge_y = 50, nudge_x = 50 ) +
    geom_point(data = StephanieRent, aes(AvgFloorPlanCost, Residents), col = "orange", fill="red", size=3.5, show.legend = FALSE) +
  labs(
    title = "Possible Housing For You",
    subtitle = "These apartments are closest to your likes",
    x="Price ($)",
    y="Max Number of Students per Semester"
  ) +
  theme_minimal()

```

<br>

These apartments are the ones that relate the most to your likes. Here are their statistics, and their information. Please note that there are many other apartments that either have more people, or are closer to campus, however those have been filtered so that only the aprtments with a monthly price of **$250 - $350** monthly, **with the most students as residents,**  and **closest to campus** show.

There are no apartments with exactly $900 rent per semester that meet all the requirements used. However, let me know if you need broader thresholds for any of these requirements (Price, Capacity, Distance), and I'll help you find which apartments fit your needs. 

<br>

```{r, warning=FALSE, message=FALSE}
Steph <- StephanieRent |> 
  select(Name, AvgFloorPlanCost, Residents, CrowFlyMetersToMC, Address, Phone) |> 
  arrange(by=CrowFlyMetersToMC) |> 
  mutate(Price = dollar(AvgFloorPlanCost),
         Distance = CrowFlyMetersToMC) |> 
  select(Name, Price, Residents, CrowFlyMetersToMC, Address, Phone)

kableExtra::kable(Steph)

```

<br>

It's your time to make a choice! I hope you enjoy your time at BYUI friend. Play with the graph below to check out the other apartments!

<br>

---
Note that this is a hypothetical situation but served to show basic analytical skills
---
