---
output: html_document
editor_options: 
  chunk_output_type: console
---
<<<<<<< HEAD
---
title: "Gapminder task 2"
date: "January 22, 2024"
gfm:
  code_fold: true
  code_tools: true
  code-line-numbers: true
execute:
  keep-md: true
---

I followed the video and used dplyr to wrangle the data

```{r}
library(tidyverse)
library(gapminder)
```

## Here is the graph needed for the assignment

```{r}

                    #Removes Kuwait and makes pop in the thousands

gapminder1 <- gapminder%>%
  
  filter(country != "Kuwait") %>% 
  
  mutate(pop = pop/100000)

                    #New Data set with weighted averages and 2.5 times pop to match the template graph

gapped <- gapminder1 %>% 
  group_by(year, continent) %>% 
  summarise(
      GDP = weighted.mean(gdpPercap, pop, na.rm=T),
      pop= pop*2.5
  )

```

```{r, fig.width=11.5}
                   #create base plot with countries

ggplot(data = gapminder1, mapping = aes(x = year, y = gdpPercap, col = continent)) +
  geom_point() +
  geom_line(aes(group=country)) +
  
                   #add weighted mean points and lines
  
  geom_point(data = gapped, mapping = aes(x = year, y = GDP, size=pop), col="black") +
  geom_line(data=gapped, mapping = aes(x=year, y=GDP, group=continent), col="black")+
  
                   #separate graphs by continent
  
  facet_wrap(vars(continent), ncol = 5)+
  
                   #labels and them
  
  labs(
    x= "Year",
    y= "GDP per capita",
    title= "Summary of population growth related to GDP per Capita by continent")+
  theme_bw()


```
=======
---
title: "Gapminder task 2"
date: "January 18, 2024"
gfm:
  code_fold: true
  code_tools: true
  code-line-numbers: true
execute:
  keep-md: true
---

I followed the video and used dplyr to wrangle the data

```{r}
library(tidyverse)
library(gapminder)
```

## Here is the graph needed for the assignment

```{r}

                    #Removes Kuwait and makes pop in the thousands

gapminder1 <- gapminder%>%
  
  filter(country != "Kuwait") %>% 
  
  mutate(pop = pop/1000000)

                    #New Data set with weighted

gapped <- gapminder1 %>% 
  group_by(year, continent) %>% 
  summarise(
      GDP = weighted.mean(gdpPercap, pop, na.rm=T),
      pop= pop)

```

```{r, fig.width=11.5}
                   #create base plot with countries

ggplot(data = gapminder1, mapping = aes(x = year, y = gdpPercap, col = continent)) +
  
  geom_point(alpha=.05, color="grey65", 
             show.legend = c( size = T,
                              color = F)) +
  
  geom_line(aes(group=country), color="grey80",
            show.legend = c( size = T,
                              color = F),
            alpha = .7) +
  
                   #add weighted mean points and lines
  
  geom_point(data = gapped, 
             mapping = aes(x = year, 
                           y = GDP, 
                           size=pop,
                           color=continent), 
             show.legend = c( size = T,
                              color = F)) +
  
  scale_y_continuous(breaks = 
                       seq(0, 50000, by = 10000),
                     labels = 
                       c("0",
                         "10k",
                         "20k",
                         "30k",
                         "40k",
                         "50k")) +
  scale_x_continuous(breaks = 
                       seq(1950, 2030, by = 20),
                     labels = 
                       c("1950",
                         "1970",
                         "1990",
                         "2010",
                         "2030")) +
  
  geom_line(data=gapped, 
            mapping = aes(x=year, 
                          y=GDP, 
                          group=continent,
                           color=continent), 
             show.legend = c( size = T,
                              color = F),
            linewidth = 1.2) +
  
                   #separate graphs by continent
  
  facet_wrap(vars(continent), ncol = 5)+
  
                   #labels and them
  
  labs(
    x= "Year",
    y= "GDP per capita",
    title= "GDP per Capita by Continent | Population Growth",
    size = "Population (M)")+
  theme_bw() +

  scale_color_manual(
    values = c(
      "chocolate", "orange", "darkseagreen", "steelblue", "skyblue2"
    )
  ) +
  
  theme(
      
    axis.text.y =
      element_text(color="grey25",size=10.5),
    
    axis.text.x = 
      element_text(color="grey25", size=10.5),
    
    axis.title = 
      element_text(color="grey15", size=12),
    
    axis.ticks = 
      element_blank(),
    
    plot.title = 
      element_text(size=14,
                   color="grey5", 
                   hjust = .5),
    
    panel.grid.major = 
      element_blank(),
    
    panel.grid.minor.x = 
      element_blank(),
    
    strip.text = 
      element_text(size = 12, 
                   face="bold", 
                   angle = 0, 
                   colour = "white"),
    
    strip.background = 
      element_rect(colour = "steelblue", 
                   fill = "steelblue"))


```



```{r}
ggplot() +
  # Plot points with size and color legends
  geom_point(
    data = gapped, 
    mapping = aes(
      x = year, 
      y = GDP, 
      size = pop,
      color = continent
    ), 
    show.legend = c(size = TRUE, color = TRUE)
  ) +
  
  scale_y_continuous(
    breaks = seq(0, 50000, by = 10000),
    labels = c("0", "10k", "20k", "30k", "40k", "50k")
  ) +
  
  geom_line(
    data = gapped, 
    mapping = aes(
      x = year, 
      y = GDP, 
      group = continent,
      color = continent
    ), 
    show.legend = c(size = TRUE, color = FALSE),
    linewidth = 1.2
  ) +
  
  scale_color_manual(
    values = c(
      "chocolate", "orange", "darkseagreen", "steelblue", "skyblue2"
    )
  ) +
  
  # Adjust the legend to make color circles larger
  guides(
    color = guide_legend(
      title = "Continent",
      override.aes = list(size = 4)  # Increase the size of color legend circles
    ),
    size = guide_legend(
      title = "Population (M)"
    )
  ) +
  
  # Labels and themes
  labs(
    x = "Year",
    y = "GDP per capita",
    title= "GDP per Capita by Continent | Population Growth",
  ) +
  theme_classic() +
  theme(
    axis.text.y = element_text(color = "grey25", size = 10.5),
    axis.text.x = element_text(color = "grey25", size = 10.5),
    axis.title = element_text(color = "grey15", size = 12),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 14, color = "grey5", hjust = 0.5),
    panel.grid.major = element_blank(),
    panel.grid.minor.x = element_blank(),
    strip.text = element_text(size = 12, face = "bold", angle = 0, colour = "white"),
    strip.background = element_rect(colour = "steelblue", fill = "steelblue")
  )
  
```

