---
title: "Week 9 Case study"
date: "March 9, 2024"
gfm:
  code_fold: true
  code_tools: true
  code-line-numbers: true
execute:
  keep-md: true
editor_options: 
  chunk_output_type: console
---



```{r, fig.width=12, fig.height=8}

library(USAboundaries)
library(tidyverse)
library(sf)
library(geofacet)

us <- us_states()
ny <- us_counties(state="NY")
permits <- buildings::permits

sing_fam_stat <- permits |> 
  filter(variable == "Single Family") |> 
  group_by(state, year) |> 
  summarise(permits=sum(value))

nyp <- permits |> 
  filter(variable == "Single Family",
         StateAbbr == "NY") |> 
  group_by(county, year) |> 
  summarise(permits=sum(value))


all_states <- us |> 
  mutate(state=as.numeric(statefp)) |> 
  inner_join(sing_fam_stat, by="state")

nyc <- ny |> 
  select(!state_name) |> 
  mutate(county=as.numeric(countyfp)) |> 
  inner_join(nyp, by="county")

ggplot(all_states, aes(x=year, y=permits/1000)) +
  
  geom_line(color='black', linewidth=1.1) +
  
  labs(
    title="Single Family Building Permits in the US",
    
    x="Year (1980 - 2010) ",
    
    y="Thousand Number of Permits") +
  
  theme_minimal() +
  
  facet_geo(
    
    ~state_abbr,
    
    grid="us_state_grid2", 
    
    strip.position="bottom") +
  
  theme(
    
    strip.text = element_text(
      
      size = 12, angle = 0, colour = "black"),
    
    strip.background = element_rect(
      
      colour = "white", fill = "white"),
    
    panel.grid.major = element_blank(),
    
    panel.grid.minor = element_blank(),
    
    axis.text.x = element_blank(), 
    
    plot.margin = margin(1, 1, 1, 1, "cm"),
    
    axis.title.y = element_text(vjust = 8),
    
    plot.title = element_text(hjust = .5))

```


```{r, fig.width=12, fig.height=15}

ggplot(nyc, aes(x=year, y=permits)) +
  
  geom_line(color='black', linewidth=1.1) +
  
  labs(
    
    title="Single Family Building Permits in New York",
    
    x="Year (1980 - 2010) ",
    
    y="Number of Permits") +
  
  theme_minimal() +
  
  facet_geo(
    
    ~name,
    
    grid="us_ny_counties_grid1",
    
    strip.position = "bottom") +
  
  theme(
    
    strip.text = element_text(
      
      size = 8, angle = 0, colour = "black"),
    
    strip.background = element_rect(
      
      colour = "white", fill = "white"),
    
    panel.grid.major = element_blank(),
    
    panel.grid.minor = element_blank(),
    
    axis.text.x = element_blank(), 
    
    plot.margin = margin(1, 1, 1, 1, "cm"),
    
    axis.title.y = element_text(vjust = 8),
    
    plot.title = element_text(hjust = .5)
  )
```


I begin to start liking geom_sf, it's really cool what you can do. The first time I used it was on a graph recreation from our world in data, they use a ton of world maps as heat maps for many different variables. I got the idea of taking out the y axis text and putting the names on the bottom from stack overflow. I chose all black to not lose focus on the data, I was trying to make the american flag in the US all states graph. But it may be unprofessional for this specific task.

Texas and Florida seem to be big builders in their areas for the last decades. New york seems average, I chose it because facet geo doesn't have all states yet so, I thought of new york. There are a couple counties in new york like saratoga that seem to be big in the construction industry

